---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: Changes in Land Use between 1990 and 2014
subtitle: https://github.com/RsMarx/Ag_Forest_Data_Final-
author: Rebecca Marx
abstract: "Experimental overview. This section should be no longer than 250 words."
fontsize: 12pt
mainfont: Times New Roman
---

<Information in these brackets are used for annotating the RMarkdown file. They will not appear in the final version of the PDF document>

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

<Setup the global options for the R chunks in your document>

<Note: set up autoreferencing for figures and tables in your document>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set your working directory
setwd("C:/Users/rsmar/OneDrive/Documents/Spring 2019/RFolder/Ag_Forest_Data_Final-")

# Load your packages
library(tidyverse)
library(ggplot2)
library(trend)
library(lubridate)
library(lsmeans)
library(nlme)
library(gridExtra)
library(knitr)
#install.packages("car")
library(car)

# Set your ggplot theme
RMtheme <- theme_bw(base_size = 10) +
  theme(plot.title = element_text(size = 16, color = "gray13", hjust = .5),
    axis.text = element_text(color = "gray13"),
    axis.title.x = element_text(size = 13),
    axis.title.y = element_text(size = 13),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
  legend.position = "bottom", legend.text = element_text(size = 12), legend.title = element_text(size = 13)) +
  theme(panel.grid.major.x = element_blank(),
  panel.grid.major.y = element_line( size=.1, color="gray99")) 
theme_set(RMtheme)

#Set up global options for R chuck in document 
knitr::opts_chunk$set(echo=TRUE, warning=FALSE)

```


```{r table, echo = FALSE}

#="\\label{tab:table}Name of Table")

```




# Research Question and Rationale

<Paragraph detailing the rationale for your analysis. What is the significant application and/or interest in this topic? Connect to environmental topic(s)/challenge(s).>

<Paragraph detailing your research question(s) and goals. What do you want to find out? Include a sentence (or a few) on the dataset you are using to answer this question - just enough to give your reader an idea of where you are going with the analysis.>

As the global population continues to grow and require more food and fuel resources, deforestation trends are expected to accelerate. Deforestation can be problematic as forests provide ecosystem services such as carbon storage, nutrient cycling, water filtration, and wildlife habitat. Agriculture is one of the most commonly sited drivers of deforestation, in addition to being a source of emissions. Given that forest and agriculture can be contending land uses, this research examines the relationship between agriculture and forest as land uses and, more broadly, explores the causes and impacts of changes in land use. 

This research looks at the trends and tradeoffs in land use across countries from 1990 – 2014. Primary questions include: 

- Is there a relationship between the percentage of land area dedicated to forest versus agricultural in countries?
- Is there a relationship between land uses (agriculture or forest) and levels of CO2, methane, and NO3 emissions?
- Is there a relationship between access to electricity or renewable electricity output and the percentage of land dedicated to forestry versus agriculture? 

The research utilizes a data set from the World Bank that has 135 environment-related variables for 264 countries. I have narrowed the environment variable down to 7 that are relevant to land use. Although the full data set dates back to 1960, I have limited to the time scope of the analysis to between 1990 and 2014 because those are the dates for which forest cover data is available. 

\newpage

# Dataset Information

## Database Information 
Data were collected from the World Bank website. More information can be found here: https://data.worldbank.org/topic/environment
Data were downloaded as a full data set called “API_6_DS2_en_csv_v2_10518155”. The csv file was saved as “WorldBank_Raw_4.8.19”. 
Variables from among the raw data set were selected to be used in the data set analysed. Only 7 of a possible 135 environment-related variables were retained in the processed data. This csv file was saved as [“WorldBank_Processed”; “WorldBank_Spread” or “WB_Spread”]. 

## Data Content Information 
Country: Originally named “Country.Name”, the Country variable includes 264 countries. 

Year: The Year variable was originally spread horizontally from 1960 – 2018. The years were gathered in to one column named “Year” and formatted as a date. 

Forest: The Forest variable was originally named “Forest area (% of land area)”. Its data range is 1990 – 2016.  

Agriculture: The Agriculture variable was originally named “Agriculture land (% of land area). Its data range is from 1961 – 2018. 

Ag.Methane: The Ag.Methane variable was originally named “Agriculture methane emissions (thousand metric tons of CO2 equivalent). Its data range is 1970 – 2008. 

Ag.NO2: The Ag.NO2 variable was originally named “Agriculture nitrous oxide emissions (thousand metric tons of CO2 equivalent). Its data range is 1970 – 2008.

CO2Emissions: The CO2Emissions variable was originally named “CO2 emissions (kt). Its data range is 1960 – 2014. 

ElectricityAccess: The ElectricityAccess variable was originally named “Access to electricity (% of total population). Data range is 1990 – 2016. 

RenewableElectricity: The RenewableElectricity variable was originally named “Renewable electricity output (% of total electricity output). Its data range is 1990 – 2015. 

## Naming conventions and file formats
The files are named according to the following convention:
`databasename_datatype_details_stage.format`, where:
Files are named according to the following naming convention: 
**databasename** refers to the database from where the data originated
**details** are additional descriptive details related to the stage of data wrangling 
**stage**refers to the stage in data management pipelines (e.g., raw or processed)
**format** is a non-proprietary file format (e.g., .csv, .txt)


## Additional Information and Support
For more information, please contact the data assembler, **Rebecca Marx** (rebecca.marx@duke.edu)


<Information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the README file for the dataset but formatted in a way that is more narrative.>

<Add a table that summarizes your data structure. This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.>

Variable     | Units     | Data Strcuture | 
------------ | --------- | -------------- |
 Cell 1      |  Cell 2   | Cells          |
 Cell 3      |  Cell 4   | Cells          |


\newpage

# Exploratory Data Analysis and Wrangling

Wrangling the data required multiple operations. First, I used the filter function to select only 9 of a possible 135 variables in the data set. Since the data set was arranged with the years going horizontally, I then had to gather the years in to one column, and create a new column called “Level” that contained the values recorded for each variable in each year. 



```{r}
World_Bank_Master <-read.csv("../Raw/WorldBank_Raw2_4.8.19.csv")

#Data Subset
World_Bank_Filter <- filter(World_Bank_Master, Indicator.Name == "Forest area (% of land area)" | Indicator.Name == "Agricultural land (% of land area)" | Indicator.Name == "Access to electricity (% of population)" | Indicator.Name == "Renewable electricity output (% of total electricity output)" | Indicator.Name == "CO2 emissions (kt)" | Indicator.Name == "Agricultural nitrous oxide emissions (thousand metric tons of CO2 equivalent)" | Indicator.Name == "Agricultural methane emissions (thousand metric tons of CO2 equivalent)")

WorldBank_Gather <- gather(World_Bank_Filter, "Year", "Level", X1960:X2018)

WorldBank_Gather <- select(WorldBank_Gather, -Indicator.Code)

WorldBank_Spread <-  spread(WorldBank_Gather, Indicator.Name, Level)

#Format as character 
WorldBank_Spread$Year <- as.character(WorldBank_Spread$Year)

#create string
WB_String <- substr(WorldBank_Spread$Year, 2, 5)

#Get rid of X in date
WorldBank_Spread$Year = WB_String

#Format as date
#WB_Fixed$Year <- as.Date(WB_Fixed$Year)
WorldBank_Spread$Year <- as.Date(WorldBank_Spread$Year, format = "%Y") #can I get it to show only the year? 

#Change column names 
names(WorldBank_Spread) <- c("Country", "Indicator.Code", "Year", "ElectricityAccess", "Agriculture", "Ag.Methane", "Ag.NO2", "CO2Emissions", "Forest", "RenewableElectricity")

#Save processed file 
#write.csv(WorldBank_Spread, row.names = FALSE, file = "../Processed/WorldBank_Processed.csv")
   
Five_Countries <- filter(WorldBank_Spread, Country == "Brazil" | Country == "Kenya" | Country == "Spain" | Country == "Indonesia" | Country == "Canada")

WB_Spread <- WorldBank_Spread %>% 
  na.exclude #check if I need this for my tests

WB_Spread2 <- WorldBank_Spread %>% 
  na.exclude

WB_Brazil <- filter(WB_Spread, Country == "Brazil")

```

```

<Include R chunks for 5+ lines of summary code (display code and output), 3+ exploratory graphs (display graphs only), and any wrangling you do to your dataset(s).> 


```
After the data was wrangled into a format that would function well with my intended R operations, I was able to begin exploring the data set. 

```{r echo=TRUE}
#5+ lines of summary 

colnames(WorldBank_Spread)
class(WorldBank_Spread$Year)
dim(WorldBank_Spread)
head(WorldBank_Spread)
summary(WorldBank_Spread)
summary(WorldBank_Spread$Agriculture)
summary(WorldBank_Spread$Forest)
summary(WorldBank_Spread$RenewableElectricty)
summary(WB_Spread$Ag.Methane)

```

Figure\ref{fig:fig1} plots agriculture and forest land covers to see if there is a noticable trend. It appears that there is a tendency for forest cover to decrease as agriculture cover increases. In other words, the scatter plot shows some correlation between the Forest and Agriculture variables. 

```{r fig1, echo=FALSE, fig.cap="\\label{fig:fig1}Percentage of Land Use for Forest and Agriculture", fig.widtch=5, fig.height=6, fig.pos='H', warning= FALSE}
suppressWarnings(expr)

#Exploratory graph 1: Agriculture vs. Forest 
AgVForest <- 
  ggplot(WorldBank_Spread) + 
  geom_point(aes(x = Agriculture, y = Forest)) +
  ggtitle("Percentage of Land Use for Forest and Agriculture") +
  ylab(expression("% Land Cover in Forest")) +
  xlab(expression("% Land Cover in Agriculture")) +
  scale_y_continuous(limits = c(0,100)) +
  scale_x_continuous(limits = c(0,100)) 
print(AgVForest)
```


A second exploratory graph (Figure\ref{fig:fig2}) plots forest levels for all 264 countries over time. The resulting graph shows that there is some movement over time, but many countries see constant levels of land use. Similarly, the plot of agriculture over time shows some movement over time, but many countries see constant levels of land use for agriculture. In later graphs I decided to reduce the number of countries included to be able to better visualized changes. 

```{r echo=FALSE, fig.cap="\\label{fig:fig2}Percentage of Land Use for Forest and Agriculture", fig.widtch=10, fig.height=10, fig.pos='H', warning= FALSE}
suppressWarnings(expr)

#Exploratory graph 2: Forest over time 
ForestTime <- 
  ggplot(WB_Spread) +
  geom_line(aes(x = Year, y = Forest, color = Country)) +
  scale_y_continuous(limits = c(0,100)) +
  ggtitle("Change in % Forest Cover Over Time") +
  ylab(expression("% Land Cover in Forest")) +
  xlab(expression("Year")) +
  theme(legend.position = "none")
#print(ForestTime)

AgricultureTime <- 
   ggplot(WB_Spread) +
  geom_line(aes(x = Year, y = Agriculture, color = Country)) +
  scale_y_continuous(limits = c(0,100)) +
    ggtitle("Change in % Agriculture Cover Over Time") +
  ylab(expression("% Land Cover in Agriculture")) +
  xlab(expression("Year")) +
  theme(legend.position = "none")
#print(AgricultureTime)

grid.arrange(ForestTime, AgricultureTime)

```


```{r}
Forest.Hist <- ggplot(WB_Spread) + 
  geom_histogram(aes(x = Forest)) +
  scale_x_continuous(limits = c(0,100))
#print(Forest.Hist)

WB_Spread2$trans_Forest <- log(WB_Spread2$Forest)

log.Forest.Hist <- ggplot(WB_Spread2) + 
  geom_histogram(aes(x = Forest)) 
#print(log.Forest.Hist)

Ag.Hist <- ggplot(WB_Spread) +
  geom_histogram(aes(x = Agriculture)) +
  scale_x_continuous(limits = c(0,100))

grid.arrange(Forest.Hist, log.Forest.Hist, Ag.Hist)

#Decided not to log transform despite the fact that the distribution was not completely normally becasue the logged results would be less intuitive. 

```


```{r echo=FALSE, fig.cap="\\label{fig:fig3}Change in Land Use from 1990 to 2014 for Five Countries", fig.widtch=10, fig.height=5, fig.pos='H', warning= FALSE}

LegendTitle1 <-"Land Use"

FiveCountries.Facet <- 
  ggplot(Five_Countries) +
  geom_line(aes(x = Year, y = Agriculture, color = "Agriculture")) + 
  geom_line(aes(x = Year, y = Forest, color = "Forest")) + 
  facet_grid(cols = vars(Country)) +
  ggtitle("Change in Land Use from 1990 to 2014") +
  ylab(expression("Land Area (%)")) +
  scale_x_date(limits = as.Date(c("1990-04-09", "2014-04-09")),
  date_breaks = "48 months", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(limits = c(0,70)) +
  scale_color_manual(LegendTitle1,values = c("goldenrod3", "darkcyan")) +
  labs(caption = "Data Source: World Bank") +
  theme(strip.background = element_rect(fill= "aquamarine3", "darkslategray")) 
print(FiveCountries.Facet)
```


Since the research attempts to decipher whether a change in forest levels is related to a change in agriculture levels, I also wanted to see if a change in the amount of land dedicated to agriculture leads to a change in various emissions levels. The following graphs explore these relationships using the 5 selected countries. 

```{r echo=FALSE, fig.cap="\\label{fig:fig4}Change in Emissions Use from 1990 to 2014 for Five Countries", fig.widtch=10, fig.height=10, fig.pos='H', warning= FALSE}

LegendTitle2 <- "Emmissions Type"

FiveCountries.Meth.NO2 <- 
  ggplot(Five_Countries) +
  geom_line(aes(x = Year, y = Ag.Methane, color = "Methane Levels")) + 
  geom_line(aes(x = Year, y = Ag.NO2, color = "NO2 Levels")) +
  facet_grid(cols = vars(Country)) +
  ggtitle("Change in Methane and NO2 Emissions from 1990 to 2014") +
  ylab(expression("Emissions (thousand metric tons)")) +
  scale_x_date(limits = as.Date(c("1990-04-09", "2014-04-09")),
  date_breaks = "48 months", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_color_manual(LegendTitle2,values = c("brown", "gray")) +
  labs(caption = "Data Source: World Bank") +
  theme(strip.background = element_rect(fill= "aquamarine3", "darkslategray")) 
#print(FiveCountries.Meth.NO2)

FiveCountries.Carbon <- 
  ggplot(Five_Countries) +
  geom_line(aes(x = Year, y = CO2Emissions, color = "CO2 Levels")) + 
  facet_grid(cols = vars(Country)) +
  ggtitle("Change in CO2 Emissions from 1990 to 2014") +
  ylab(expression("CO2 Emissions (thousand metric tons)")) +
  scale_x_date(limits = as.Date(c("1990-04-09", "2014-04-09")),
  date_breaks = "48 months", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_color_manual(values = c("Orange")) +
  labs(caption = "Data Source: World Bank") +
  theme(strip.background = element_rect(fill= "aquamarine3", "darkslategray")) 
#print(FiveCountries.Carbon)

grid.arrange(FiveCountries.Meth.NO2, FiveCountries.Carbon)

```


```


```
```{r echo=FALSE, fig.cap="\\label{fig:fig5} Renewable Electricity and Electricity Access from 1990 to 2014 for Five Countries", fig.widtch=10, fig.height=10, fig.pos='H', warning= FALSE}

facet_fill_color <- c("red", "green", "blue", "yellow", "orange")

FiveCountries.RE <- 
  ggplot(Five_Countries) +
  geom_line(aes(x = Year, y = RenewableElectricity, color = "Renewable Energy Output")) + 
  geom_line(aes(x = Year, y = ElectricityAccess, color = "Population with Electricty Access")) +
  facet_grid(cols = vars(Country)) +
  ggtitle("Changes in Renewable Electricity and Electricity Accessfrom 1990 to 2014") +
  ylab(expression("Percentage of Total")) +
  scale_x_date(limits = as.Date(c("1990-04-09", "2014-04-09")),
  date_breaks = "48 months", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_color_manual(values = c("purple", "blue")) +
  labs(caption = "Data Source: World Bank") +
  theme(strip.background = element_rect(fill = "aquamarine3", "darkslategray")) 
#print(FiveCountries.RE)

Test <- 
  ggplot(Five_Countries) +
  geom_line(aes(x = Year, y = RenewableElectricity, color = "Renewable Energy Output")) + 
  geom_line(aes(x = Year, y = ElectricityAccess, color = "Population with Electricty Access")) +
  facet_grid(cols = vars(Country)) + 
  ggtitle("Changes in Renewable Electricity and Electricity Accessfrom 1990 to 2014") +
  ylab(expression("Percentage of Total")) +
  scale_x_date(limits = as.Date(c("1990-04-09", "2014-04-09")),
  date_breaks = "48 months", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_color_manual(values = c("purple", "blue")) +
  labs(caption = "Data Source: World Bank") +
  theme(strip.background = element_rect(fill = "aquamarine3", "darkslategray")) 
print(Test)


```


<Include text sections to accompany these R chunks to explain the reasoning behind your workflow, and the rationale for your approach.>


\newpage

# Analysis
<Include R chunks for 3+ statistical tests (display code and output) and 3+ final visualization graphs (display graphs only).>





```{r echo=TRUE}
#Statistical Test 1: How has forest changed over time 
Forest.Time1 <- gls(data = WB_Spread, 
                    Forest ~ Year,
                    method = "REML")
summary(Forest.Time1)
#Forest decreasing over time (-.00424, p-value 0.00)
#AIC 49977.96

Forest.Time2 <- lme(data = WB_Spread,
                Forest ~ Year,
                random = ~1| Country,
                method = "REML")
summary(Forest.Time2)
#Forest decreasing over time (-.00421, p-value 0.00)
#AIC 49431.95

#AIC for the random effect model is lower, so used this going forward. 

Logit_Test <- glm(data = WB_Spread,
             Forest ~ Year)
summary(Logit_Test)
#The results were not much different than using gls, and I was not able to account for variance. 
```

A fixed effects model was used to see how land percetnage of forest area accross the whoel data set changes over time. The results show that, on average, forest decreases by -.00485% each year. Was not worride about seasonality of the data because there is only one data point per year. As such, I did not calculate autocorrelations. I did add country as a random effect. 

```{r echo=TRUE}
For.Ag.1 <- gls(data = WB_Spread, 
                    Forest ~ Year + Agriculture,
                    method = "REML")
summary(For.Ag.1)
#AIC: 49908 

For.Ag.2 <-  lme(data = WB_Spread,
              Forest ~ Year + Agriculture,
              random = ~1| Country)
summary(For.Ag.2)
#AIC: 49413

cor.test(formula = ~ Forest + Agriculture,
         data = WB_Spread)
#There is some negative correlation between Agriculture and Forest with a 95% confidence interval of -.16 and -.10. 

```

Results of first set of tests: 

I am interested in whether access to electricity or the levels of renewable energy produced in countries have a relationship with deforestation. This is becasue a lack of electricity could drive the use of forest biomass products as an energy source. Electricity access appears to have a negative impact on forest (-.333) which may be because electricity is needed for some agricultural operations, or electricity is being produced via biofuels, the production of which might require more forest to be converted to agricultural land. ( \ref{fig:fig3})

With that theory in mind, I was interested in whether there might be a relationship between forested land and renewable energy because energy produced from biomass is considered to be renewable. However, the result of my gls revealed that a 1 unit increase in renewable energy is associated with a .18% increase in forest land, meaning the production of biofule is likely not a main driver of deforestation. 

```{r echo=TRUE}
#Looking at Electricity Accesss and Renewable Energy as a Drive of the Ag. / Forest Tradeoff

#Add Electricity 
Forest.Ag.Elec <- gls(data = WB_Spread, 
                    Forest ~ Year + Agriculture + ElectricityAccess,
                    method = "REML")
summary(Forest.Ag.Elec)
#AIC: 49823.17

Forest.Ag.Elec2 <- lme(data = WB_Spread, 
                    Forest ~ Year + Agriculture + ElectricityAccess,
                    random = ~1| Country)
summary(Forest.Ag.Elec2)
#AIC: 49390

car::vif(Forest.Ag.Elec2)
cor(Agriculture, ElectricityAccess)


#Electricity only 
Forest.Elec <- lme(data = WB_Spread, 
                    Forest ~ Year + ElectricityAccess,
                    random = ~1 | Country)
summary(Forest.Elec)


Forest.RE <- lme(data = WB_Spread, 
                    Forest ~ Year + RenewableElectricity,
                    random = ~1 | Country)
summary(Forest.RE)


```

As a second statistical test I used Pettit’s test, a nonparametric test that determines if there is a shift in the central tendency of the time series. It also determines at what point in time the changepoint occurs. I used this to see if there was a point in time in particular where these seemed to be a noticeable change in forest cover. I applied the test to Brazil which graphically appears to have change points to see if 1) there was a change poing for agriculture and forest and 2) to see if the change points occured in the same year for each. 


```{r echo=TRUE}
#Statistical Test 2: Pettitts Test: Looking at Change Points in the Data

#Statistical Test 2: Any change points for full forest data?
pettitt.test(WB_Spread$Forest)
#Probable change point at time 1428 which doesn't exist.


#Changes in Brazil Forest
pettitt.test(WB_Brazil$Forest)
#9 (of 19): 1998; p-value .048 (significant at the .05 level)

pettitt.test(WB_Brazil$Forest[9:19])

#Changes in Brazil Agriculture 
#5 (9+5 = 14: 2003); p-value .0023852 (significant at the .01 level)

pettitt.test(WB_Brazil$Forest[14:19])
#3 14 + 3 = 17: 2006 (p-value .29: not significant)

pettitt.test(WB_Brazil$Agriculture)
#9 (of 19): 1998; p-value .002386 (significant at the .01 level)

pettitt.test(WB_Brazil$Agriculture[9:19])
#5 (9+5 = 14: 2003); p-value .04852 (significant at the .05 level

pettitt.test(WB_Brazil$Agriculture[14:19])
#2 (14 +2 = 16) p-value .8487 (not significant)


```

Pettitt's applied to a single country (Brazil) initially detects a change point in Forest and Agriculture in the same year. The change for both was in place 9, which is the year 1998. The significant change points for agriculture and forest in Brazil occured in the same years (1998 and 2003), furhter supporting the notion that there is a relationship between the two. 

[Make a graph of Brazil that shows the change points]


The next series of tests address the question: Is there a relationship between land uses (agriculture or forest) and levels of CO2, methane, and NO3 emissions?

```{r echo=TRUE}
#Statistical Test 3.1: Methane Emmissions

AgMethane <- gls(data = WB_Spread,
                 Ag.Methane ~ Year + Agriculture,
                 method = "REML")
summary(AgMethane)

ForMethane <- gls(data = WB_Spread,
                 Ag.Methane ~ Year + Forest,
                 method = "REML")
summary(ForMethane)


Ag.For.Methane <- gls(data = WB_Spread,
                 Ag.Methane ~ Year + Agriculture + Forest,
                 method = "REML")
summary(Ag.For.Methane)

car::vif(Ag.For.Methane)


#Statistica Test 3.2: Forest & Methane Emmissions
ForestMethane <- gls(data = WB_Spread, 
                  Ag.Methane ~ Year + Forest,
                  method = "REML")
summary(ForestMethane)

#Statistical Test 3.3: Interaction of Forest and Agriculture on Methane Emissions #These are probably correleated. Run a VIF Test. 
Int.Methane <-  gls(data = WB_Spread,
                    Ag.Methane ~ Year + Forest * Agriculture,
                    method = "REML")
summary(Int.Methane)
```

As the percetage of land under agriculture increases by 1, methane emissions increase by 6.57 thousand metric tons. 

As Forest land increases by 1%, methane emissions increase by .035 thoushand metric tons (or 35 metirc tons), meaning that deforestation is not strongly related to methane emissions. 

However, if the interaction of forest and agriculture land cover is considered, an increase of 1 leads to an increase of 98.86 thousand metric tons of methane. 

```{r echo=TRUE}
#Carbon Emissions Tests 

#Test 4.1: Forest and Carbon
Forest.CO2 <- gls(data = WB_Spread, 
                  CO2Emissions ~ Year + Forest,
                  method = "REML")
summary(Forest.CO2)

#Test 4.2: Agriculture and Carbon 
Ag.CO2 <- gls(data = WB_Spread,
                 CO2Emissions ~ Year + Agriculture,
                 method = "REML")
summary(Ag.CO2)

#Test 4.3: Interaction of Forest & Agriculture on Carbon 
Ag.For.CO2 <- gls(data = WB_Spread,
                    CO2Emissions ~ Year + Forest + Agriculture,
                    method = "REML")
summary(Ag.For.CO2)

car::vif(Ag.For.Methane)

```
1% increase in forest leads to a -235.25 decrease in CO2. 
1% increase in Ag. leads to an increase of 1493 kt of carbon. 
Interaction: How do you interpret? 


<Include text sections to accompany these R chunks to explain the reasoning behind your workflow, rationale for your approach, and the justification of meeting or failing to meet assumptions of tests.>


\newpage

# Summary and Conclusions
<Summarize your major findings from your analyses. What conclusions do you draw from your findings? Make sure to apply this to a broader application for the research question you have answered.>



